An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This work has been partially funded by CoMeDiC ANR-15-CE40-0006 research grant.}

### 2020-01-06 jmiv.tex

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage[utf8]{inputenc}
\usepackage{graphicx,subfig}
\usepackage{amstext,amsmath,amssymb,bm,bbm,mathtools}
\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\usepackage[export]{adjustbox}
\usepackage{xcolor}

@@

\begin{document}
%
\title{A\HT{n Elastica-driven} Digital Curve Evolution Model for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}
  \titlerunning{Digital Curve Evolution by Elastica Energy}
  %\title{Digital Curvature Evolution Models for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}

\author{Daniel Antunes\inst{1}
Jacques-Olivier Lachaud\inst{1}
Hugues Talbot\inst{2}}
%
\authorrunning{D. Antunes et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Universit{\'e} Savoie Mont Blanc, LAMA, UMR CNRS 5127, F-73376, France
\email{daniel.martins-antunes, jacques-olivier.lachaud@univ-savoie.fr} \and
CentraleSupelec Inria \'equipe OPIS, Universit\'e Paris-Saclay, 9 rue Joliot-Curie, F-91190 Gif-sur-Yvette \\
\email{hugues.talbot@centralesupelec.fr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  \HT{Geometric priors have been shown to be useful in image segmentation to regularize results. For example, the classical
    Mumford-Shah functional uses regions perimeter as prior. This has inspired much research in the last few decades,
    with classical approaches like the Rudin-Osher-Fatemi and most graph-cut formulations, which all use a weighted or
    binary perimeter prior. It has been observed that this prior is not suitable in many applications, for example for
    segmenting thin objects or some textures, which may have high perimeter/surface ratio. Mumford observed that an
    interesting prior for natural objects is the Euler Elastical model, which involves the squared curvature. In other
    areas of science, researchers have noticed that some physical binarization processes, like emulsion unmixing can be
    well-approximated by curvature-related flow like the Willmore flow. 
    However, curvature-related flows are not easy to compute because curvature is difficult to estimate accurately, and the
    underlying optimisation processes are not convex. In this article, we propose to formulate a digital flow that
    approximates a curvature-related flow using a multigrid-convergent curvature estimator, within a discrete
    variational framework. We also present an application of this model as a post-processing step in a segmentation
    framework.}
\keywords{Multigrid convergence  \and Digital estimator \and Curvature \and Shape Optimization \and Image Segmentation.}
\end{abstract}
%
%
%
\setcounter{footnote}{0}
\section{Introduction} %% \todo{(To be extended)}}

@@

In fact, it is still a very challenging task to inject curvature in
the context of image segmentation.  
\HT{State-of-art methods are in practice} difficult to optimize and \HT{do not scale}
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
\HT{often} unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods.

Segmentation methods using Elastica or Willmore energies \HT{generally
reduce to} computing the corresponding flow, i.e. following its
gradient. Well-founded level-set methods involving Willmore energy
\cite{droske2004level} have been used in medical image segmentation
\cite{lim2012introducing}. Another level-set formulation similar to
the Chan-Vese model \HT{was} proposed in \cite{zhu2013image}. Willmore
flow can also be carried out by phase-field models (see
\cite{bretin2015phase}) but they are less suited to image segmentation
since interfaces are blurred in such models. Threshold dynamics can
also be considered for such energies \cite{esedoglu2008threshold} and
has been proposed for image desocclusion \cite{esedoglu2005threshold}.

In this paper we are interested in purely discrete variational segmentation models involving an elastica
energy. Recently, new estimators of curvature along digital contours have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the \HT{desirable} multigrid convergence property. This
motivates us to \HT{propose} models in which they can be used successfully.

\HT{In the remainder, we} investigate the use of the digital integral invariant
curvature estimator \cite{coeurjolly13integral} in a discrete
variational segmentation framework. More precisely, we show how to
incorporate it in a digital flow minimizing its squared
curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO
algorithm \cite{rother07qpbo}. \HT{Our solution does not} compute a global
solution to the whole domain, but our approach leads to a digital flow
similar to continuous \HT{Elastica}, as shown by our experiments. We
present an application of the model as a post-processing step in a
segmentation framework, and demonstrate how it can improve standard
results. The code is freely available on
github\footnote{https://www.github.com/danoan/BTools}.

\begin{itemize}
\item{Geometric properties and regularization}
\item{The role of curvature}

\item{What are the difficulties in use curvatue as a regularizer (point out convex relaxation approaches) and how a digital approach may help us }
\item{Outline: We present two models that illustrate how digital estimators can be used to minimize the elastica energy on digital sets.  }
\item{\todo{Add related works : geometric priors, binary segmentation, other digital curve evolution models}}
\item{\todo{Talk about Willmore flow with phase field or with threshold dynamics}}

\item{\textbf{Convex relaxation}}
  \begin{itemize}
  \item{A convex lower semicontinuous approximation of Eulers elastica energy \cite{bredies15convex}}
  \item{Introducing Total Curvature for Image Processing \cite{goldluecke11totalcurvature}}
  \end{itemize}

\item{\textbf{Constraint optimization}}
  \begin{itemize}
  \item{Curvature regularity for region-based image segmentation and inpainting: A linear programming relaxation \cite{schoenemann09linear} }	
  \item{A Fast Algorithm for Euler's Elastica Model Using Augmented Lagrangian Method \cite{tai11elastica} }		
  %%\item{The elastic ratio: Introducing curvature into ratio-based image segmentation \cite{schoenemann2011elastic}
  %%}		%% DONE
  \item{Curvature Regularization for Curves and Surfaces in a Global Optimization Framework \cite{strandmark11globalframework} }
  \end{itemize}

\item{\textbf{Discrete approaches}}
  \begin{itemize}
  \item{Fast global optimization of curvature \cite{zehiry10fast}	}
  \item{Efficient Squared Curvature \cite{nieuwenhuis14efficient} 	}
  \item{Contrast driven {E}lastica for image segmentation \cite{el2016contrast} }	
  \end{itemize}

\end{itemize}

\textit{Outline}. Section two reviews the concept of multigrid
convergence and highlights its importance for the definition of
digital estimators. Next, we describe two convergent estimators used
in this paper, one for tangent and the other for curvature. They are
used in the optimization model and in the definition of the digital
elastica. Section three describes a local combinatorial optimization model
suitable for different curvature estimators. Section four describes the proposed 
curvature evolution model along with several illustrations of digital flows. Section five
explains how to use the evolution model as a post-processing step in
an image segmentation framework.

\section{Multigrid convergent estimators \todo{(Mostly copy/paste)}}

The goal of this section is to  introduce the concept of multigrid convergenc and its potential when analyzing digital images with variational models involving geometric quantities. More precisely we wish to delineate shapes within digital images with some priors related to their unknown continuous geometry. Of course we only observe shape digitizations within images.
%% \begin{itemize}
%% 	\item{Why to use multigrid convergent estimators}
%% 	\item{Multigrid convergence for local estimators}
%% 	\item{Reference to examples of multigrid digital estimators(for curvature, length, tangent)}
%% 	\item{Define the elastica energy estimator}
%% 	\item{\test{(Jaco) see if two opposite II evaluations give convergent curvature}}
%% \end{itemize}

%% \hrulefill

@@

\begin{figure}[!ht]
\center
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h05/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h05/summary-flow.eps}
	}%
		\caption{Local combinatorial optimization process results for the square and flower shapes. Shapes displayed at every $2$ iterations for the square and $7$ iterations for the flower.}	
		\label{fig:local-comb-square-results}
\end{figure}

The running time of algorithm \ref{alg:local-search} is summarized in
table \ref{tab:summary-local-comb-rtime}. All the experiments in this paper were executed on a five core $3.4Ghz$ cpu and the number of pixels in the triangle, square and flower shapes are respectively $841,1867,521$ for grid step $h=1.0$. Although its use in practical applications is limited, we
demonstrate that digital estimators are effective in their measurements
and the flows evolve as expected. We observe that it is a
complete digital approach, and we do not suffer from discretization
and rounding problems, a common issue in continuous models.
Furthermore we have checked that this approach works
indifferently with Integral Invariant curvature estimator and
Maximal Digital Circular Arc curvature estimator. So the convergence
of the digital curvature estimator seems to be the cornerstone to
get a digital curve behaving like a continuous elastica.  In the
next section, we explore a more efficient approach.

\begin{figure}
\begin{minipage}[b]{0.5\textwidth}
\center
\includegraphics[scale=0.35]{images/local_search/square-flower.eps}
\caption{Elastica computation along iterations of algorithm \ref{alg:local-search}}
\label{fig:plot-elastica-local-search}
\end{minipage}\hspace{40pt}%
\begin{minipage}[b]{0.45\textwidth}
\center
\captionsetup{type=table}
\begin{tabular}{|l|c|c|}
\hline
& $h=1.0$ & $h=0.5$\\
\hline
Square & 47s (5s/it) & 260s (12s/it)\\
\hline
Flower & 1235s (23s/it)  & 8516s (52s/it)\\
\hline
\end{tabular}
\caption{Running times for the local combinatorial optimization algorithm with $k=1,N=20$. Four threads were used.}
\label{tab:summary-local-comb-rtime}
\end{minipage}
\end{figure}

@@

In figure \ref{fig:m1-square-flow} we see some results for the DCE
algorithm with $m=1$. We observe a global movement towards roundness,
but with several artifacts. We minimize the effects of a jaggy
boundary by setting $\alpha > 0$. Nonetheless, a higher estimation
ball radius creates unstable shapes. In fact, the estimator is very
sensitive in regions of low squared curvature, and it is precisely in
those regions that spurious pixels are created.

\begin{figure}[!ht]
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-nolength/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-length0.15/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r5/summary_flow.eps}
	}%
\end{minipage}%
\caption{The algorithm is very sensitive to the little variations of the estimator, which are particularly important in regions of low squared curvature. Artifacts are somewhat reduced with a length penalization but increases if we use a higher ball radius. }
\label{fig:m1-square-flow}
\end{figure}

\subsection{A more stable model.}
In the previous section we noticed that the algorithm produces shapes
with many artifacts due to the small uncertainties of the estimator
along regions of low squared curvature. We argue that, evaluating the
estimation ball along farther ring sets we avoid those sensitive areas
by focusing the optimization process only on regions with highest
squared curvature value.

@@

\begin{figure}[!ht]
\center
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=1$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l1/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=2$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l2/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=3$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l3/summary_flow.eps}
	}%
\end{minipage}%
\caption{By positioning the estimation ball on farther rings, we minimize artifacts creation. \label{fig:mx-square-flow}}
\end{figure}

In our experiments, the best results are obtained by executing DCE
algorithm with $m$ equal to $r$, where $r$ is the estimation ball
radius (see figures \ref{fig:mx-square-flow},
\ref{fig:mx-flow-gs-radius-effect} ). In plot
\ref{fig:mx-elastica-plots}, we observe how the elastica energy evolves along
the iterations. If the radius is too large, the elastica energy may increase after some steps. That is what happens for the
triangle, in which the flow converges to a single point.

\begin{figure}[!ht]
\center				
		\setcounter{subfigure}{-3}
		\subfloat[$r=3,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r3-h0.5/summary_flow.eps}
		}%
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r3-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r3-h0.5/summary_flow.eps}}%
		}

		\setcounter{subfigure}{-2}
		\subfloat[$r=5,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r5-h0.5/summary_flow.eps}
		}%1
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r5-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r5-h0.5/summary_flow.eps}}%
		}				
\caption{The choice of radius impacts the flow. In the figures, the flow ceases to evolve for all shapes  when $r=3$ (a). In figure (b), for $r=5$, the triangle evolves to a single point, while the others stop  in an intermediate shape, as in (a). }
\label{fig:mx-flow-gs-radius-effect}
\end{figure}

\begin{figure}[!ht]
\center
\includegraphics[scale=0.5]{images/flow/elastica-energy-plot/r5h05.eps}
\caption{Elastica computation along the iterations for different shapes. If radius is too large, elastica energy may increase after a certain number of iterations. }
\label{fig:mx-elastica-plots}
\end{figure}

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
%\documentclass[runningheads]{llncs}
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%
\usepackage[utf8]{inputenc}
\usepackage{graphicx,subfig}
\usepackage{amstext,amsmath,amssymb,bm,bbm,mathtools}
\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
\usepackage[export]{adjustbox}
\usepackage[dvipsnames]{xcolor}

@@

\begin{document}
%
\title{An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}
\titlerunning{Digital Curve Evolution by Elastica Energy}
  %\title{Digital Curvature Evolution Models for Image Segmentation\thanks{This  work has  been  partially  funded by CoMeDiC ANR-15-CE40-0006 research grant.}}

\author{Daniel Antunes%\inst{1}
\and Jacques-Olivier Lachaud%\inst{1}
\and Hugues Talbot%\inst{2}
}
%
\authorrunning{D. Antunes et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{%Universit{\'e} Savoie Mont Blanc, LAMA, UMR CNRS 5127, F-73376, France
  Daniel Antunes \at
  Univ. Grenoble Alpes, Univ. Savoie Mont Blanc, CNRS, LAMA, 73000 Chambéry, France \\
  \email{daniel.martins-antunes@univ-smb.fr} \and
  Jacques-Olivier~Lachaud \at  Univ. Grenoble Alpes, Univ. Savoie Mont Blanc, CNRS, LAMA, 73000 Chambéry, France \\
  \email{jacques-olivier.lachaud@univ-smb.fr} \and
  Hugues Talbot \at CentraleSupelec, Inria, Universit\'e Paris-Saclay, 9 rue Joliot-Curie, F-91190 Gif-sur-Yvette \\
  \email{hugues.talbot@centralesupelec.fr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
  Geometric priors have been shown to be useful in image segmentation to regularize results. For example, the classical
    Mumford-Shah functional uses region perimeter as prior. This has inspired much research in the last few decades,
    with classical approaches like the Rudin-Osher-Fatemi and most graph-cut formulations, which all use a weighted or
    binary perimeter prior. It has been observed that this prior is not suitable in many applications, for example for
    segmenting thin objects or some textures, which may have high perimeter/surface ratio. Mumford observed that an
    interesting prior for natural objects is the Euler Elastical model, which involves the squared curvature. In other
    areas of science, researchers have noticed that some physical binarization processes, like emulsion unmixing can be
    well-approximated by curvature-related flow like the Willmore flow. 
    However, curvature-related flows are not easy to compute because curvature is difficult to estimate accurately, and the
    underlying optimisation processes are not convex. In this article, we propose to formulate a digital flow that
    approximates an Elastica-related flow using a multigrid-convergent curvature estimator, within a discrete
    variational framework. We also present an application of this model as a post-processing step to a segmentation
    framework.
\keywords{Multigrid convergence  \and Digital estimator \and Curvature \and Shape Optimization \and Image Segmentation.}
\end{abstract}
%
%
%
\setcounter{footnote}{0}
\section{Introduction} %% \todo{(To be extended)}}

@@

In fact, it is still a very challenging task to inject curvature in
the context of image segmentation.  
\HT{State-of-art methods are in practice} difficult to optimize and \HT{do not scale}
=======
known about the shape of interest. Length penalization is a
well-behaved, general purpose regularizer and many models in the
literature make use of it, from active contours
\cite{casseles97geodesic} to level-set formulations
\cite{malladi1995image,malladi1995shape}. Discrete graph-based
variational models have been particularly succesful to incorporate
length penalization as a penalizer, while keeping the ability to
extract a global optimum \cite{boykov01graphcut,appleton05geodesic}.

However length regularization shows limitations when segmenting small
or thin and elongated objects, as it tends to shrink solutions or
yields disconnected solutions. Such drawbacks can be problematic in
image segmentation, image restoration or image inpainting. It is thus
rather natural to consider curvature (especially squared curvature) as
a potential regularizer. Energies involving both length and squared
curvature are often called the {\em Elastica} model (that dates back
to Euler). The {\em Willmore energy} is its $n$-dimensional
extension. These were brought to attention in computer vision by
Mumford \cite{mumford1994elastica}. A similar concept is also present
in the second-order regularizer of the original snake model
\cite{kass1988snakes}.Indeed, an explicit curvature term was often
employed in early level-set methods,
like~\cite{malladi1995image,malladi1995shape,ballester01filljoint} for
segmentation or inpainting. However, Casselles {\em et al.}
in~\cite{casseles97geodesic} observed that these terms as employed
were not geometric, in the sense that they depend on the
discretization parameters.

\subsection{Existing works}
The use of curvature for surface regularization in the case of the
Willmore energy was studied in~\cite{bobenko2005discrete}. Although
results look promising, authors used an inexact linearized version of
the square curvature at every step.

One of the first successful uses of geometric, squared curvature in
image processing is the inpainting algorithm described in
\cite{masnou98inpainting}, where authors evaluate the absolute
curvature along the level lines of a simply-connected image to
reconstruct its occluded parts. The non-intersection property of level
lines induces the construction of an efficient dynamic programming
algorithm. The curvature energy is nevertheless only coarsely
approximated. We may quote \cite{chan02elasticainpainting} as another
geometric inpainting method involving Elastica, which is optimized with
level-sets.

In image segmentation, Zehiry and Grady have shown that injecting
curvature as a regularizer can help recover thin and elongated objects
\cite{zehiry10fast}. Similarly Schoenemann {\em et al.}
\cite{schoenemann09linear} have considered the Elastica energy in
their ratio-based image segmentation
approach. In~\cite{schoenemann2011elastic}, Schoenemann~{\em et al.}
extend ratio-cut approaches to segmentation with a curvature
term. Their global optimization framework shows the considerable
advantages of using such regularizer in common binary image
segmentation tasks.  However the time complexity of their algorithm
is prohibitive, even if curvature is again coarsely
approximated.  In a more recent work, Lim {\em et al.} propose an
edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set
formulation. They observe improved results compared
to~\cite{casseles97geodesic} and other perimeter-based priors, but
observe that a shape prior and a close-enough initialization must also be
provided to achieve good results.

In fact, it is still a very challenging task to efficiently handle
curvature in the context of image segmentation.  State-of-art methods
are in practice difficult to optimize and do not scale easily
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
often unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods.

Segmentation methods using Elastica or Willmore energies can be
formulated via the computation of the corresponding flow,
i.e. following its gradient. Well-founded level-set methods involving
Willmore energy \cite{droske2004level} have been used in medical image
segmentation \cite{lim2012introducing}. Another level-set formulation
similar to the Chan-Vese model was proposed in
\cite{zhu2013image}. Willmore flow can also be carried out by
phase-field models (see \cite{bretin2015phase}) but they are less
suited to image segmentation since interfaces are blurred in such
models. Threshold dynamics can also be considered for such energies
\cite{esedoglu2008threshold} and has been proposed for image
desocclusion \cite{esedoglu2005threshold}. However, flow-based methods
will always deliver a local minimum, which heavily depend on
initialization for good results. A similar approach is to formulate
the Elastica as a regularization in a variational framework and use
operator splitting to optimize it. Tai {\em et
  al.}~\cite{tai11elastica} and more recently Deng~{\em et
  al.}~\cite{deng2019new} have proposed this approach, which is
well-suited to inverse problem solving. The problem remains non-convex
and the nature of the converged iterates are not studied, i.e. how
close they are to a global optimum.

In contrast, researchers have sought convex relaxation of curvature-related
formulation. In~\cite{goldluecke11totalcurvature}, Goldluecke and Cremers introduce the {\em Total Curvature}, based
on the Menger-Melnikov curvature of a measure. The formulation is non-convex but can be relaxed into a well-correlated
but non-tight convex one. The relaxed formulation can be used to solve inverse problems including segmentation. However computing a solution is still expensive, requiring signifiant parallelisation efforts with a GPU to achieve
  acceptable computing times. In~\cite{bredies15convex}, Bredies {\em et al.} study a convex approximation of the
  Elastica energy, using a lifting scheme. Thanks to the lifting, it is not restricted to binary sets. It shows
  promising results and can be used for image restoration in addition to inpainting and segmentation. The scheme is
  however neither tight nor exact, and required 4 dimensions for a 2D approximation. In addition, it does not carry over
to $nD$. It also remains computationally expensive.

Discrete approaches have been used to tackle the Elastica
difficulties. As noted before, \cite{schoenemann09linear} have
linearized a mesh-based approach, which was expanded upon
by~\cite{strandmark11globalframework}, while \cite{zehiry10fast} have
used a binary solver for a non-submodular approach on a regular square
grid of pixels, which was expanded
in~\cite{el2016contrast}. In~\cite{nieuwenhuis14efficient} , a
specific solver is developed to yield a better solution to an
extension of the formulation in~\cite{zehiry10fast}.

\subsection{Motivation}
In spite of the multiplicity of existing approaches, it seems none is
currently completely satisfactory. While the sought solution is
expected to be very smooth based on geometrical expectations, the
mathematical properties of the related operators make this smoothness
elusive. Non-convexity and the presence of high-order terms make
Elastica difficult to optimize and to approximate, let alone compute
efficiently. The computed solutions in the literature are rarely
evaluated, and the complexity of the proposed solution is often high.

%%\item{What are the difficulties in use curvature as a regularizer (point out convex relaxation approaches) and how a digital approach may help us }
%%\item{Outline: We present two models that illustrate how digital estimators can be used to minimize the elastica energy on digital sets.  }
%%\item{\todo{Add related works : geometric priors, binary segmentation, other digital curve evolution models}}
%%\item{\todo{Talk about Willmore flow with phase field or with threshold dynamics}}

Consequently, we are interested in studying purely discrete
variational segmentation models involving an Elastica
energy. Recently, new estimators of curvature along digital contours
have been proposed
\cite{roussillon11mdca,coeurjolly13integral,schindele17mdca} with the
desirable multigrid convergence property. This motivates us to propose
models in which they can be used successfully.

In this paper, we propose to investigate the use of the digital
integral invariant curvature estimator \cite{coeurjolly13integral} in
a discrete variational segmentation framework. More precisely, we show
how to incorporate it in a digital flow minimizing its squared
curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO
algorithm \cite{rother07qpbo}. Our solution does not compute a global
solution to the whole domain but a global solution to a band
surrounding a given shape. Our approach thus leads to a digital flow
similar to continuous Elastica, as shown by our experiments. We
present an application of this model as a post-processing step in a
segmentation framework, and demonstrate how it can improve standard
results. This paper is an extended version of a previous work that
appeared in \cite{antunes19}. We have included a deeper discussion on
our optimization method and we present a more thorough comparison of
our segmentation method with two other related works. Moreover, we
have included a new local combinatorial optimization approach for
elastica minimization, which show that our discrete model is
well-founded. Finally, the code produced in this paper is freely
available on github.\footnote{https://www.github.com/danoan/BTools}

\subsection{Outline}
Section~2 reviews the concept of multigrid convergence and highlights
its importance for the definition of digital estimators. We describe
two convergent estimators used in this paper, one that approaches the
length of elementary discrete contour elements, and the other that
approaches the curvature of a discrete contour. They are used in the
optimization model and in the definition of the digital
elastica. Section~3 describes a local combinatorial optimization model
suitable for several curvature estimators, that presents interesting
results but is too costly in practice. Section~4 describes the
proposed Elastica-driven curvature evolution model along with several
illustrations of digital flows. Section~5 explains how to use this
evolution model as a post-processing step in an image segmentation
framework and compares it to two related approaches. We conclude and
draw some perspectives to this work in Section~6.

\section{Multigrid convergent estimators}
Our objective is to delineate shapes within digital images with some priors related to continuous geometry. The goal of
this section is to introduce the concept of multigrid convergence and its potential when analyzing digital images with
variational models involving geometric quantities, with the constraint that only digitized shapes are observed.

@@

\begin{figure}[!h]
\center
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/square/h05/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=1.0\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h1/summary-flow.eps}
	}%
	\hspace{10pt}
	\subfloat[h=0.5\label{}]{%
	\includegraphics[scale=0.125]{images/local_search/flower/h05/summary-flow.eps}
	}%
		\caption{Local combinatorial optimization process results for the square and flower shapes. Shapes displayed at every $2$ iterations for the square and $7$ iterations for the flower.}	
		\label{fig:local-comb-square-results}
\end{figure}

The running time of algorithm \ref{alg:local-search} is summarized in table \ref{tab:summary-local-comb-rtime}. All the
experiments in this paper were executed on a five-core $3.4Ghz$ CPU and the number of pixels in the triangle, square and
flower shapes are respectively $841,1867,521$ for grid step $h=1.0$. Although its use in practical applications is
limited, we demonstrate that digital estimators are effective in their measurements and the flows evolve as expected. We
observe that it is a complete digital approach, and we do not suffer from discretization and rounding problems, a common
issue in continuous models.  Furthermore we have checked that this approach works indifferently with Integral Invariant
curvature estimator and Maximal Digital Circular Arc curvature estimator. So the convergence of the digital curvature
estimator seems to be the cornerstone to get a digital curve behaving like a continuous Elastica.  In the next section,
we explore a more efficient approach.

\begin{figure}[!h]
\begin{minipage}[b]{0.5\textwidth}
\center
\includegraphics[scale=0.35]{images/local_search/square-flower.eps}
\caption{Elastica computation along iterations of algorithm \ref{alg:local-search}}
\label{fig:plot-elastica-local-search}
\end{minipage}\hspace{40pt}%
\begin{minipage}[b]{0.45\textwidth}
\center
\captionsetup{type=table}
\begin{tabular}{|l|c|c|}
\hline
& $h=1.0$ & $h=0.5$\\
\hline
Square & 47s (5s/it) & 260s (12s/it)\\
\hline
Flower & 1235s (23s/it)  & 8516s (52s/it)\\
\hline
\end{tabular}
\caption{Running times for the local combinatorial optimization algorithm with $k=1,N=20$. Four threads were used.}
\label{tab:summary-local-comb-rtime}
\end{minipage}
\end{figure}

@@

In figure \ref{fig:m1-square-flow} we see some results for the DCE algorithm with $m=1$. We observe a global evolution
towards rounder shapes, but with several artifacts. We minimize the effects of a jaggy boundary by setting $\alpha >
0$. Nonetheless, a higher estimation ball radius creates unstable shapes. In fact, the estimator is very sensitive in
regions of low squared curvature, and it is precisely in those regions that spurious pixels are created.

\begin{figure}[!h]
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-nolength/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=3, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r3-length0.15/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, \alpha=0.15$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/length-radius-effect/triangle-r5/summary_flow.eps}
	}%
\end{minipage}%
\caption{The algorithm is very sensitive to the little variations of the estimator, which are particularly important in regions of low squared curvature. Artifacts are somewhat reduced with a length penalization but increases if we use a higher ball radius. }
\label{fig:m1-square-flow}
\end{figure}

\subsection{A more stable model.}
In the previous section we noticed that the algorithm produces shapes with many artifacts due to the small uncertainties
of the estimator along regions of low squared curvature. We argue that, evaluating the estimation ball along outer
ring sets we avoid those sensitive areas by focusing the optimization process only on regions with highest squared
curvature value.

@@

\begin{figure}[!h]
\center
\center
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=1$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l1/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=2$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l2/summary_flow.eps}
	}%
\end{minipage}%
\begin{minipage}[b]{0.33\textwidth}
	\subfloat[$r=5, m=3$\label{}]{%
	\includegraphics[scale=0.2]{images/flow/farther-rings/square-r5-l3/summary_flow.eps}
	}%
\end{minipage}%
\caption{By positioning the estimation ball on farther rings, we minimize artifacts creation. \label{fig:mx-square-flow}}
\end{figure}

{In our experiments, the best results are obtained by executing DCE algorithm with $m$ equal to $r$, where $r$ is
  the estimation ball radius (see figure \ref{fig:mx-square-flow} ). We observe that Elastica may increase after some
  iterations if chosen radius is too large, as in the case of the triangle in figure \ref{fig:mx-flow-gs-radius-effect}
  in which the flow converges to a single point. We conjecture that an appropriated value for the radius should be given
  by the shape reach.  The produced flow has no difficulties in handling changes on topology, and it presents different
  speeds for regions with low and high curvature values, as illustrated in figure
  \ref{fig:mx-speed-variation-hole-filling}.

\begin{figure}[hp!]
\center				
		\setcounter{subfigure}{-3}
		\subfloat[$r=3,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r3-h0.5/summary_flow.eps}
		}%
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r3-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r3-h0.5/summary_flow.eps}}%
		}

		\setcounter{subfigure}{-2}
		\subfloat[$r=5,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r5-h0.5/summary_flow.eps}
		}%1
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r5-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r5-h0.5/summary_flow.eps}}%
		}\\
		\subfloat[Elastica evaluation]{%
		\includegraphics[scale=0.5]{images/flow/elastica-energy-plot/r5h05.eps}						
		}
\caption{The choice of radius impacts the flow. In the figures, the flow ceases to evolve for all shapes  when $r=3$ (a). In figure (b), for $r=5$, the triangle evolves to a single point, while the others stop  in an intermediate shape, as in (a). In figure (c), we observe that for a given choice of radius, the elastica may increase after a certain number of iterations. }
\label{fig:mx-flow-gs-radius-effect}
\end{figure}
